---
output:
 html_document:
    fig_width: 5
    fig_height: 2.9
    fig.align: "center"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, 
                      warning=FALSE, results='hide', fig.keep='all', 
                      cache = TRUE, autodep = TRUE)
# echo = TRUE print the R code
```

# Data


# Objective



# Setup

## Load data

Make sure you've downloaded the [`movies_merged`](https://s3.amazonaws.com/content.udacity-data.com/courses/gt-cs6242/project/movies_merged) file and it is in the current working directory. Now load it into memory:

```{r}
setwd("/Users//Desktop//PR1")
load('movies_merged')
cat("Dataset has", dim(movies_merged)[1], "rows and", dim(movies_merged)[2], "columns", end="\n", file="")
```

This creates an object of the same name (`movies_merged`). For convenience, you can copy it to `df` and start using it:

```{r}
df = movies_merged
cat("Column names:", end="\n", file="")
colnames(df)
```

## Load R packages


```{r}
install.packages("reshape")
install.packages("reshape2")
install.packages("plyr")
install.packages("tm")
```


```{r}
library(ggplot2)
library(GGally)
library(reshape)
library(reshape2)
library(plyr)
```

**Non-standard packages used**: 
```{r}
# used the below for multiplot
# http://www.peterhaschke.com/r/2013/04/24/MultiPlot.html
source("http://peterhaschke.com/Code/multiplot.R")

# for tm package
# install tm package
# install.packages("tm")
library(tm)

# install stringr package
# install.packages("stringr")
library(stringr)
```

# Tasks

## 1. Remove non-movie rows


```{r}
df <- df[df$Type == "movie",]
dim(df)
```

**Q**: How many rows are left after removal? 

**A**: 40000

## 2. Process `Runtime` column


```{r}

# replace anything that is hard coded as "N/A" by 0
df$Runtime[df$Runtime=="N/A"] = 0

# replace anything that is NA by 0
df$Runtime[is.na(df$Runtime)] = 0

# split into different parts of the string
df1 = strsplit(df$Runtime, " ", perl = TRUE)

for (i in 1:length(df1)) {
  # handle when there is no unit on second position i.e just the number
  if (is.na(df1[[i]][2])) {
    df$Runtime[i] = as.numeric(df1[[i]][1])
  } 
  # for h only, check the first part to be h and the 4th part to be NA
  else if (((df1[[i]][2] == "h") & (is.na(df1[[i]][4])))){
    df$Runtime[i] = (60*as.numeric(df1[[i]][1]))
  }
  # handle when there is h and min
  else if ((df1[[i]][2] == "h") & (df1[[i]][4] == "min")){
    df$Runtime[i] = ((60*as.numeric(df1[[i]][1])) + (as.numeric(df1[[i]][3])))
  }
  # handle when there is only min
  else if (df1[[i]][2] == "min") {
    df$Runtime[i] = (as.numeric(df1[[i]][1]))
  }
  else {
    # ignore
  }
  # convert all to numeric one last time
  df$Runtime = as.numeric(df$Runtime)
}
```


```{r, warning=FALSE, message=FALSE, fig.retina = 1}

# create plots
# scatter plot
ggplot(df, aes(Year, Runtime)) + geom_point() + 
  theme(axis.text.x =
          element_text(size  = 9,
                       hjust = 1)) +
  scale_alpha_continuous(breaks = seq(1850, 2050, by=10)) +
  ggtitle("Figure 1") +
  xlab("Year") + 
  ylab("Total movie run time in a year [min]")
```


```{r, warning=FALSE, message=FALSE, fig.retina = 1}
# create histogram of the run time
ggplot(df, aes(Runtime)) + geom_histogram() + xlab("Total run time in a year [min]") +
  xlim(c(0,200)) + ggtitle("Figure 2")

#after few testing set the max limit to 200 for histogram
```


```{r, warning=FALSE, message=FALSE, fig.retina = 1}
# Total Runtime vs. year
df2 = aggregate(Runtime ~ Year, df, sum)

ggplot(df2, aes(Year, Runtime)) + 
  geom_point() + 
  theme(axis.text.x =
          element_text(size  = 9,
                       angle = 90,
                       hjust = 1)) +
  scale_x_continuous(breaks = seq(1850, 2017, by=10)) + ggtitle("Figure 3") +
  xlab("Year") + 
  ylab("Total movie run time in a year [min]")
```


```{r, warning=FALSE, message=FALSE, fig.retina = 1}
# Runtime vs. budget

# scatter plot
ggplot(df, aes(Runtime, Budget))+ geom_point() + 
  theme(axis.text.x =
          element_text(size  = 9,
                       hjust = 1)) +
  ggtitle("Figure 4") + xlim(0,350) +
  xlab("Runtime [min]") + ylab("Budget")
```


```{r}

# replace all "N/A" with NA on Genre
df$Genre[df$Genre=="N/A"] = NA

# replace ", " with " "
df$Genre = gsub(", ", " ", df$Genre)
# https://piazza.com/class/j6gt7ycx6nk145?cid=542
# replace "-" with ""
df$Genre = gsub("-", "", df$Genre)
# convert genre column to binary
genre_corp = Corpus(VectorSource(df$Genre))
genre_dtm = DocumentTermMatrix(genre_corp)
df = data.frame(df, as.matrix(genre_dtm))
# delete genre column
df = subset(df, select = -Genre)
```

Plot the relative proportions of movies having the top 10 most common genres.

```{r}

#sum the values in each genre column and sort in decreasing order
# after visual inspection of data frame found that the new genre columns spans from coumn 39 to 68
df_genre_col_sum = sort(colSums(df[,39:66]), decreasing = T)[1:10]

list = names(df_genre_col_sum)
# print (list) 

df_genre_888 = df[list]

df_99999 = colSums(Filter(is.numeric, df_genre_888))

df_99999 = data.frame(df_99999)

df_99999 = data.frame("Genre" = list, "Sum" = df_99999[,1])

df_99999 = data.frame(df_99999, "Relative" = df_99999[,2]/sum(df_99999$Sum))

ggplot(df_99999, aes(Genre, Relative)) + geom_bar(stat = "identity") + 
  theme(axis.text.x =
          element_text(size  = 9,
                       angle = 90,
                       hjust = 1)) +
  xlab("Genre") + 
  ylab("Relative proportions") +
   ggtitle("Figure 5")
```


```{r, warning=FALSE, message=FALSE, fig.retina = 1}

# select the required column in new dataframe
df_genre_10 = df[,c("Runtime", list)]
# get the totla of run time for each genre
# length(list)
vec_sum = c()
vec_genre = c()

# dim(df_genre_10) # 11 col but the genre names are in 2 to 11
for (i in 2:11) {
  vec_genre = c(vec_genre, list[i-1])
  vec_sum1 = sum(df_genre_10[df_genre_10[c(i)] == 1,]$Runtime)
  vec_sum = c(vec_sum, vec_sum1)
}

df_genre_total_runtime = data.frame("Runtime" = vec_sum, "Genre" = vec_genre)

ggplot(df_genre_total_runtime, aes(Genre, Runtime)) + geom_bar(stat = "identity") + 
  theme(axis.text.x =
          element_text(size  = 9,
                       angle = 90,
                       hjust = 1)) +
  xlab("Genre") + 
  ylab("Total movie run time for the genre [min]") +
   ggtitle("Figure 6") 
```


```{r, warning=FALSE, message=FALSE, fig.retina = 1}
# subset the dataframe by each genre type and plot as bx plot then combine all box plot into one 
drama = ggplot(subset(df_genre_10, drama ==1), aes(drama, Runtime)) + geom_boxplot()
comedy = ggplot(subset(df_genre_10, comedy ==1), aes(comedy, Runtime)) + geom_boxplot()
short = ggplot(subset(df_genre_10, short ==1), aes(short, Runtime)) + geom_boxplot()
romance = ggplot(subset(df_genre_10, romance ==1), aes(romance, Runtime)) + geom_boxplot()
action = ggplot(subset(df_genre_10, action ==1), aes(action, Runtime)) + geom_boxplot()
crime = ggplot(subset(df_genre_10, crime ==1), aes(crime, Runtime)) + geom_boxplot()
thriller = ggplot(subset(df_genre_10, thriller ==1), aes(thriller, Runtime)) + geom_boxplot()
documentary = ggplot(subset(df_genre_10, documentary ==1), aes(documentary, Runtime)) + geom_boxplot()
adventure = ggplot(subset(df_genre_10, adventure ==1), aes(adventure, Runtime)) + geom_boxplot()
animation = ggplot(subset(df_genre_10, animation ==1), aes(animation, Runtime)) + geom_boxplot()

# need to run below code as per
# http://www.peterhaschke.com/r/2013/04/24/MultiPlot.html
source("http://peterhaschke.com/Code/multiplot.R")

multiplot(drama, comedy, short, romance, action, crime, thriller, documentary, adventure, animation, cols=5) 
```

```{r}

# subset data frame that contains only Year, Date, Released and Gross

#check NA in all columns
sum(is.na(df$Year)) # 0
sum(is.na(df$Date)) # 35442 If we use date to remove we will end up removing almost all data
sum(is.na(df$Released)) # 4949
sum(is.na(df$Gross)) # 35442
# extract year from released year
out = strsplit(as.character(df$Released), '-')
# add to data frame
df = data.frame(df, do.call(rbind, out))
names(df)[names(df) == 'X1'] = 'Release_Year'
names(df)[names(df) == 'X2'] = 'Release_Month'
names(df)[names(df) == 'X3'] = 'Release_Day'
# typeof(df$Year[1])
# typeof(df$Release_Year[1])
# Release year has 4949 NA, convert them all to the date from the Year column
df$Release_Year[is.na(df$Release_Year)] = df$Year
# count number of rows that match on year and release_year column
# count = df[df$Year == df$Release_Year,] # 29376 rows are equal and gives a new data frame count with 29376 rows
count = df$Year == df$Release_Year # logical vector TRUE if equal and FALSE if not equal
# sum(count) # 29367
count = ifelse(!is.na(df$Gross), TRUE, count) # if the gross has NA return TRUE so as to not remove all NA gross rows
# sum(count)
df = df[count, ]
```


## 5. Explore `Gross` revenue

```{r, warning=FALSE, message=FALSE, fig.retina = 1}

# rm(df_genre_col_sum, list_2, df_genre_all, vec_sum_2, vec_genre_2, vec_sum1, df_genre_col_sum_2)
df_genre_col_sum_2 = sort(colSums(df[,39:66]), decreasing = T)

list_2 = names(df_genre_col_sum_2)
# print (list_2) 

# select the required column in new dataframe
df_genre_all = df[,c("Gross", list_2)]
vec_sum_2 = c()
vec_genre_2 = c()
#genre is listed under col 2 to 29
for (i in 2:29) {
  vec_genre_2 = c(vec_genre_2, list_2[i-1])
  vec_sum1 = sum(df_genre_all[df_genre_all[c(i)] == 1,]$Gross, na.rm = T)
  vec_sum_2 = c(vec_sum_2, vec_sum1)
}

df_genre_total_gross = data.frame("Gross" = vec_sum_2, "Genre" = vec_genre_2)

ggplot(df_genre_total_gross, aes(Genre, Gross)) + geom_bar(stat = "identity") + 
  theme(axis.text.x =
          element_text(size  = 9,
                       angle = 90,
                       hjust = 1)) +
  xlab("Genre") + 
  ylab("Gross Revenue") + 
  ggtitle("Figure 7") 
```


```{r, warning=FALSE, message=FALSE, fig.retina = 1}
# gross rev vs budget
ggplot(df, aes(Budget, Gross)) + geom_point() + 
  theme(axis.text.x =
          element_text(size  = ,
                       hjust = 1)) +
  xlab("Budget") + ylab("Gross Revenue") +  
  ggtitle("Figure 8") 
```


```{r, warning=FALSE, message=FALSE, fig.retina = 1}
# gross rev vs run time
ggplot(df, aes(Runtime, Gross)) + geom_point() + 
  theme(axis.text.x =
          element_text(size  = 9,
                       hjust = 1)) +
  xlim(c(0,350)) +
  xlab("Runtime [min]") + ylab("Gross Revenue") +  
  ggtitle("Figure 9") 
```


```{r, warning=FALSE, message=FALSE, fig.retina = 1}

# only select Gross and Month column as its easier to use
# https://stackoverflow.com/questions/1660124/how-to-sum-a-variable-by-group

vec_month = c("Jan", "Feb", "Mar",
              "Apr", "May", "June", "July",
              "Aug", "Sep", "Oct", "Nov", "Dec")

# find sum of gross for all months
df1 = aggregate(Gross ~ Release_Month, df, sum)
# add name of months to the dataframe
df1 = data.frame(df1, "Month" = vec_month)
# order so that months are in order not in alphabetical order
df1$Month = factor(df1$Month, order = T, levels= vec_month)

ggplot(df1, aes(Month, Gross)) + geom_bar(stat = "identity") + 
  theme(axis.text.x =
          element_text(size  = 9,
                       angle = 90,
                       hjust = 1)) +
  xlab("Month") + 
  ylab("Gross Revenue") + ggtitle("Figure 10") 
```


```{r}

# replace NA by 0
df$Awards[is.na(df$Awards)] = 0
# replace hard coded N/A by 0
df$Awards[df$Awards=="N/A"] = 0

# rm(vec_win, vec_nom)
vec_win = c()
vec_nom = c()

for (i in 1:dim(df)[1]) {
  string = df$Awards[i]
  string = gsub("[.]","",string) # remove the dot [.]
  string = as.list(strsplit(string, ' ')[[1]]) # split the string 
    if (length(string)==2) { # case of 1 win or 1 nomination
      if(string[2] == "win" || string[2] == "wins") {
        vec_win = c(vec_win, as.numeric(string[1])) 
        vec_nom = c(vec_nom, 0) # when you assign win you also need to assign nom
      } else if(string[2] == "nomination" || string[2] == "nominations") {
        vec_nom = c(vec_nom, as.numeric(string[1]))
        vec_win = c(vec_win, 0)
      } else { #
        vec_win = c(vec_win, 0)
        vec_nom = c(vec_nom, 0)
      }
    } 
    else if (length(string)==5) { # 1 win $ 5 nominations
      if(string[2] == "win" || string[2] == "wins") {
        vec_win = c(vec_win, as.numeric(string[1]))
        vec_nom = c(vec_nom, 0)
      } else if(string[5] == "nomination" || string[5] == "nominations") {
        vec_nom = c(vec_nom, as.numeric(string[4]))
        #vec_win = c(vec_win, 0)
      } else { #
        vec_win = c(vec_win, 0)
        vec_nom = c(vec_nom, 0)
      }
    }
    else { # ignore these "Nominated for 1 Oscar. Another 13 wins & 15 nominations.""
      vec_win = c(vec_win, 0)
      vec_nom = c(vec_nom, 0)
    }
}
# Add win and Nominations columns

df = data.frame(df, "Win" = vec_win, "Nomination" = vec_nom)
```


```{r}
sum(!df$Awards==0)
# 11491
```


```{r, warning=FALSE, message=FALSE, fig.retina = 1}

ggplot(df, aes(Win, Gross)) + geom_point() + 
  theme(axis.text.x =
          element_text(size  = 9,
                       angle = 90,
                       hjust = 1)) +
  xlab("Win") + 
  ylab("Gross Revenue") + ggtitle("Figure 11") 
```


```{r, warning=FALSE, message=FALSE, fig.retina = 1}
# df6 = aggregate(Gross ~ Win, df, sum)
# remove 0 win data
ggplot(aggregate(Gross ~ Win, df, sum), aes(Win, Gross)) + geom_point() + 
  theme(axis.text.x =
          element_text(size  = 9,
                       angle = 90,
                       hjust = 1)) +
  xlab("Win") + xlim(1,42) +
  ylab("Gross Revenue") + ggtitle("Figure 12")
```


```{r, warning=FALSE, message=FALSE, fig.retina = 1}
ggplot(df, aes(Nomination, Gross)) + geom_point() + 
  theme(axis.text.x =
          element_text(size  = 9,
                       angle = 90,
                       hjust = 1)) +
  xlab("Nomination") + 
  ylab("Gross Revenue") + ggtitle("Figure 13")
```


```{r, warning=FALSE, message=FALSE, fig.retina = 1}
# df7 = aggregate(Gross ~ Nomination, df, sum)
# did not consider with 0 nom
ggplot(aggregate(Gross ~ Nomination, df, sum), aes(Nomination, Gross)) + geom_point() + 
  theme(axis.text.x =
          element_text(size  = 9,
                       angle = 90,
                       hjust = 1)) +
  xlab("Nomination") + xlim(1,16) + ylim (0, 10000000000) +
  ylab("Gross Revenue") + ggtitle("Figure 14")
```



## 7. Movie ratings from IMDb and Rotten Tomatoes


```{r, warning=FALSE, message = FALSE, fig.retina=1}

# colnames(df)
# with meta data column, required to change cardinality_threshold to 98
#ggpairs(df, columns = c(14, 15, 16, 19, 21, 22, 23, 24, 26, 27, 28), cardinality_threshold = 98) + ggtitle("Figure 14: with metadata") 

#w/o meta data
#ggpairs(df, columns = c(15, 16, 19, 21, 22, 23, 24, 26, 27, 28)) + ggtitle("Figure 15: w/o metadata") 

```


```{r}
# cor among imdb and tomato user ratings
ggpairs(df, columns = c(15, 16, 26, 27, 28)) + ggtitle("Fig 14: cor among imdbRating, imdbVotes, tomatoMeterUser, tomatoUserRating and tomatoUserReviews") +
   ggtitle("Figure 15")
```


```{r}
# cor among Gross, imdbRating and imdbVotes
ggpairs(df, columns = c(37, 15, 16)) +  ggtitle("Figure 16: cor among Gross, imdbRating and imdbVotes")
```


```{r}
# cor among various tomato ratings
ggpairs(df, columns = c(19, 21, 22)) + ggtitle("Fig 17: cor among tomatoMeter, tomatoRating and tomatoReviews")
```


```{r}
# cor among various tomato ratings and gross
ggpairs(df, columns = c(37, 19, 21, 22)) + ggtitle("Fig 18: cor among Gross, tomatoMeter, tomatoRating and tomatoReviews")
```


```{r}
# cor among tomato user rating
ggpairs(df, columns = c(26, 27, 28)) + ggtitle("Fig 19: cor among tomatoMeterUser, tomatoUserRating and tomatoUserReviews")
```


```{r}
# cor among tomato user rating and Gross
ggpairs(df, columns = c(37, 26, 27, 28)) + ggtitle("Fig 20: cor among Gross, tomatoMeterUser, tomatoUserRating and tomatoUserReviews")
```


## 8. Ratings and awards

```{r}

# take three rating; imdbRating, tomatoRatings and tomatoUserRating
ggpairs(df, columns = c(70, 15, 21, 27)) + ggtitle("Fig 21: Cor among award win, imdbRating, tomatoRating and tomatoUserRating")
```


```{r}
ggpairs(df, columns = c(71, 15, 21, 27)) + ggtitle("Fig 22: Cor among award nominations, imdbRating, tomatoRating and tomatoUserRating")
```


## 9. Expected insights


```{r}
df_rating = df[,c("Rated", "Gross", "Budget")]
# replace anything that is hard coded as "N/A" or UNRATED or NOT RATED by NA
df_rating$Rated[df_rating$Rated=="N/A"] = NA
df_rating$Rated[tolower(df_rating$Rated)=="unrated"] = NA
df_rating$Rated[tolower(df$Rated)=="not rated"] = NA
df_rating$Rated[tolower(df$Rated)=="passed"] = NA
df_rating$Rated[tolower(df$Rated)=="approved"] = NA

df_rr = df_rating[complete.cases(df_rating[, "Rated"]),]

df_r1 = aggregate(. ~ Rated, df_rr, sum)

melted = melt(df_r1, id.vars="Rated")
```


```{r}

df_r2 = count(df_rr, "Rated")

ggplot(df_r2, aes(Rated, freq)) + geom_point() + 
  theme(axis.text.x =
          element_text(size  = 9,
                       angle = 90,
                       hjust = 1)) +
  xlab("Movie rating") + 
  ylab("Frequency") +
   ggtitle("Figure 23")
```


```{r}
ggplot() + geom_line(data=melted, aes(x=Rated, y=value, group=variable, colour=variable)) +
   ggtitle("Figure 24")
```

```{r}
ggpairs(df, columns = c(36, 37)) + ggtitle("Cor between gross and domestic gross") +
   ggtitle("Figure 25")
```


## 10. Unexpected insight

```{r}

ggpairs(df, columns = c(35, 15, 21, 27)) + ggtitle("Fig 26: Cor among movie budget, imdbRating, tomatoRating and tomatoUserRating")
```


